---
title: "Ensemble motif enhanced network inference prediction"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
This GitHub page is created to support the original [report](https://github.com/Rrtk2/MSB1014-Network_Biology-Project/Project/Report), requested by the course MSB1014 (Network Biology).

## Programming language
 - R
 - Additional: Cytoscape

#### Script block 1: INSTALL PACKAGES
This block installs all packages which are required during this algorithm. It iteratively checks if they are installed, installs if needed, and load them afterward. 
```{r message=FALSE, warning = FALSE}
	# Install packages if needed and load, or just load packages
	if (!requireNamespace("BiocManager", quietly = TRUE))
	  install.packages("BiocManager", ask = F)

	# Insert all packages in requiredpackages
	requiredpackages <-
	  c("GRENITS",
		"igraph",
		"corrr",
		"GENIE3",
		"parmigene",
		"RCy3")
		
	for (i in requiredpackages) {
		if (!requireNamespace(i, quietly = TRUE))
			BiocManager::install(i, ask = F, dependencies = TRUE)
		require(as.character(i), character.only = TRUE)
		print(i)
	}
```

#### Script block 2: SETTINGS
In this script some parameters can be changed, these are located in this block. This way changing parameters is easy and robust.
```{r}
	options(stringsAsFactors 	= F)
	motifsize 					= 3
	DataFolderLocation 			= "/Github/MSB1014-Network_Biology-Project/Project/Notebook/" 	# Must be in format "/location...DatalocationFolder/"
```



#### Script block 3: Set working directories
The working directory can be set for multiple devices, load the data and create a timestamp folder.
```{}
	# Data_dir should contain the data file!
	if(Sys.info()[6]=="RR"){
		data_dir <- paste("D:",DataFolderLocation,sep="")  # Rick's link (laptop)
	}else if(Sys.info()[6]=="Admin"){
		data_dir <- paste("D:/Games",DataFolderLocation,sep="")  # Rick's other link (desktop)
	}

	# Get folder and move to one level above, then make new folder containing timestamp
	setwd(data_dir)
	
	data <- read.delim("insilico-data.txt", row.names=1)
	set.seed(1) # for reproducibility of results

	# Make the "Results timestamp" folder
	setwd('..')
	root_dir <- getwd()
	results_dir <- paste(root_dir,"/Results ",gsub(format(Sys.time(), "%x_%H.%M.%S"), pattern = "/",replacement = "-"),sep="")
	dir.create(file.path(results_dir))
	setwd(results_dir)
```


#### Script block 4: Load true graph
A literal description of the true graph.
```{}
	graph.True = graph_from_literal(ErbB3.P--+Ras.GTP,ErbB3.P--+Gab1.P,ErbB3.P--+ErbB1.P,ErbB3.P--+PIP2,ErbB3.P--+SHC.P,
	ErbB3.P--+ErbB2.P,ErbB3.P--+PIP3,ERK.P.P--+ErbB3.P,ERK.P.P--+Gab1.P,ERK.P.P--+SOS.P,ERK.P.P--+ErbB1.P,ERK.P.P--+ErbB2.P,
	ERK.P.P--+ERK.P,Ras.GTP--+ErbB3.P,Ras.GTP--+ErbB1.P,Ras.GTP--+ErbB4.P,Ras.GTP--+SHC.P,Ras.GTP--+RAF.P,Ras.GTP--+ErbB2.P,
	MEK.P.P--+ERK.P.P,MEK.P.P--+MEK.P,MEK.P.P--+ERK.P,Gab1.P--+ErbB3.P,Gab1.P--+Ras.GTP,Gab1.P--+ErbB1.P,Gab1.P--+ErbB4.P,
	Gab1.P--+PIP2,Gab1.P--+ErbB2.P,Gab1.P--+PIP3,ErbB1.P--+ErbB3.P,ErbB1.P--+Ras.GTP,ErbB1.P--+Gab1.P,ErbB1.P--+SOS.P,
	ErbB1.P--+ErbB4.P,ErbB1.P--+PIP2,ErbB1.P--+SHC.P,ErbB1.P--+ErbB2.P,ErbB1.P--+PIP3,AKT.P--+AKT.P.P,ErbB4.P--+Ras.GTP,
	ErbB4.P--+Gab1.P,ErbB4.P--+ErbB1.P,ErbB4.P--+PIP2,ErbB4.P--+SHC.P,ErbB4.P--+ErbB2.P,ErbB4.P--+PIP3,PIP2--+Gab1.P,
	PIP2--+ErbB4.P,PIP2--+ErbB2.P,PIP2--+PIP3,MEK.P--+MEK.P.P,SHC.P--+ErbB3.P,SHC.P--+Ras.GTP,SHC.P--+SOS.P,SHC.P--+ErbB1.P,
	SHC.P--+ErbB4.P,SHC.P--+ErbB2.P,RAF.P--+Ras.GTP,RAF.P--+MEK.P.P,RAF.P--+MEK.P,ErbB2.P--+ErbB3.P,ErbB2.P--+Ras.GTP,
	ErbB2.P--+Gab1.P,ErbB2.P--+ErbB1.P,ErbB2.P--+ErbB4.P,ErbB2.P--+PIP2,ErbB2.P--+SHC.P,ErbB2.P--+PIP3,ERK.P--+ERK.P.P,
	PIP3--+AKT.P,PIP3--+PIP2,PIP3--+AKT.P.P,AKT.P.P--+AKT.P,AKT.P.P--+RAF.P,unconnected=dummy1,dummy2,dummy3)

	#nodes:
	V(graph.True)
	length(V(graph.True))

	#links:
	E(graph.True)
	length(E(graph.True))
```

#### Script block 5: function fGetMotifsFromGraph
This is a user made function, which does most of the heavy lifting of the true analysis. It creates several images and prompts metrics.
```{}
	# Get motifs fromgraph and plot the motifs with frequency

	fGetMotifsFromGraph = function(graph.name, motifsize = 3, directed=TRUE, amountRandomNets = 1000, betaFactor = 1.3, HardTH = 3,runCytoscape=FALSE){
		# motifsize = 3 # motif size: 3 or 4
		# directed = TRUE # directed 
		# amountRandomNets = 1000 # amount of random nets
		# betaFactor = 1.3 # The probability distribution in 'amountRandomNets' networks is risen to the power of beta, to cut on motifs
		# HardTH = 3 # minimum amount of ocurrences
		# runCytoscape = T/F # True for network generation in cytoscape
		
		# create folder
		root_dir <- getwd()
		results_dir <- paste(root_dir,"/",graph.name,sep="")
		dir.create(file.path(results_dir))
		setwd(results_dir)

		graph = get(graph.name)
		graphInfer = graph # used later
		#if(exists("randomNetMotifsTotal")){rm(randomNetMotifsTotal)}
		#if(exists("endresult")){rm("endresult")}
		# Find network motifs in the graph "graph":
		mygraphmotifs <- graph.motifs(graph, motifsize)
		# Find which motifs occur:

		for (i in 1:length(mygraphmotifs)){
			if(is.na(mygraphmotifs[i])){next}
			

			motif <- mygraphmotifs[i]
			if (motif > 0){ # There are some occurrences of this motif
			
				# print(i)
				# Find out what the motif looks like:
				motifgraph <- graph.isocreate(size=motifsize, number=i-1, directed=directed)
				#edges <- E(motifgraph)

				png(file = paste("Motif ", i-1, ".png", sep = ""), 
					width = 2400/2, height = 2400/2, pointsize=25)
					plot(motifgraph)
					text(-0.5,-1.2,paste("This motif(",i-1,") occurs ",motif," times:",sep=""))
				dev.off()
				
				print(paste("Motif",i-1,"occurs",motif,"times."))
				#cat(edges)
			   
			}
		}
		
		print(mygraphmotifs)
		mygraphmotifs[is.na(mygraphmotifs)]=0
		
		# start statistical testing
		graph = delete.vertices(graph, degree(graph)==0)
		
		meanInDegree = mean(degree(graph,mode = "in"))
		meanOutDegree = mean(degree(graph,mode = "out"))
		
		# start random net generation; create distribution
		for(p in 1:amountRandomNets){
			#g <- sample_pa(n=length(V(graph)),m=round(length(E(graph))/length(V(graph))),directed = directed)
			g <- random.graph.game(n = length(V(graph)),p.or.m =length(V(graph)),directed = T,type = "gnm")
			randomNetMotifs = graph.motifs(g, motifsize)
			randomNetMotifs[is.na(randomNetMotifs)]=0
			
			if(!exists("randomNetMotifsTotal")){
				randomNetMotifsTotal = randomNetMotifs
				}else{
				randomNetMotifsTotal = randomNetMotifsTotal + randomNetMotifs
			}
		}
		
		randomNetMotifsTotal = randomNetMotifsTotal/amountRandomNets
		
		#####################
		# Which occured more than random:
		impMotifIndex = which(mygraphmotifs>(randomNetMotifsTotal^betaFactor)&mygraphmotifs>HardTH)-1 #@RRR include TH morethan 3
		if(length(impMotifIndex) == 0 ){
			print("no significant motifs")
			}else{
			print(impMotifIndex)

			# get locations
				for(i in 1:length(impMotifIndex)){
				pattern = graph.isocreate(size=motifsize, number=impMotifIndex[i], directed=directed)
				iso <- subgraph_isomorphisms(pattern, graph)      # takes a while
				motifsZ <- lapply(iso, function (x) { induced_subgraph(graph, x) })

				for( p in 2:length(motifsZ)){
					if(!exists("a")){a=motifsZ[[1]]}else{a=new_g}

					b=motifsZ[[p]]

					#V(a)$name <- V(a)$label
					#V(b)$name <- V(b)$label

					attrs <- rbind(as_data_frame(a, "vertices"), as_data_frame(b, "vertices")) %>% unique()
					el <- rbind(as_data_frame(a), as_data_frame(b))
					#@RRR check if this really is one motif
					new_g <- graph_from_data_frame(el, directed = TRUE, vertices = attrs)
				}

				rm(a)
				rm(b)

				png(file = paste("Complex net from motif ", impMotifIndex[i], ".png", sep = ""), 
					width = 2400/2, height = 2400/2, pointsize=25)
					plot(new_g)
					text(-0.5,-1.2,paste("Assembled network contains ",length(motifsZ)," motifs(",impMotifIndex[i],").",sep=""))
					title("Complex")
				dev.off()

				png(file = paste("Simple net from motif ", impMotifIndex[i], ".png", sep = ""), 
					width = 2400/2, height = 2400/2, pointsize=25)
					plot(simplify(new_g))
					text(-0.5,-1.2,paste("Assembled network contains ",length(motifsZ)," motifs(",impMotifIndex[i],").",sep=""))
					title("Simplified")
				dev.off()

				# check below!...

				if(!exists("endresult")){a=new_g}else{a=endresult}

				b=new_g

				#V(a)$name <- V(a)$label
				#V(b)$name <- V(b)$label

				attrs <- rbind(as_data_frame(a, "vertices"), as_data_frame(b, "vertices")) %>% unique()
				el <- rbind(as_data_frame(a), as_data_frame(b))

				endresult <- graph_from_data_frame(el, directed = TRUE, vertices = attrs)

				rm(a)
				rm(b)
			}

			png(file = paste("Endresult complex net.png", sep = ""), 
				width = 2400, height = 2400, pointsize=25)
				plot((endresult))
				#text(-0.5,-1.2,paste("Assembled network contains ",length(motifsZ)," motifs(",impMotifIndex[i],").",sep=""))
				#title("Simplified")
			dev.off()

			png(file = paste("Endresult simple net.png", sep = ""), 
				width = 2400, height = 2400, pointsize=25)
				plot(simplify(endresult))
				#text(-0.5,-1.2,paste("Assembled network contains ",length(motifsZ)," motifs(",impMotifIndex[i],").",sep=""))
				#title("Simplified")
			dev.off()				

			
			# calculate difference and check the overlap
			
			## NO motif filter
			# missed this amount of edges:
			NoE = E(graphInfer)%in%E(graph.True)
			print(paste(sum(NoE),"/",length(NoE)," of total ",length(E(graph.True))," edges correct (inferred net)",sep=""))
			
			# missed this amount of nodes:
			NoN = V(graphInfer)%in%V(graph.True)
			print(paste(sum(NoN),"/",length(NoN)," of total ",length(V(graph.True))," nodes correct (inferred net)",sep=""))
			
			## motif filter
			# missed this amount of edges:
			YesE = E(simplify(endresult))%in%E(graph.True)
			print(paste(sum(YesE),"/",length(YesE)," of total ",length(E(graph.True))," edges correct (motif filter)",sep=""))
			
			# missed this amount of nodes:
			YesN = V(simplify(endresult))%in%V(graph.True)
			print(paste(sum(YesN),"/",length(YesN)," of total ",length(V(graph.True))," nodes correct (motif filter)",sep=""))
			# ...to this!
			
			# Cytoscape!
			if(runCytoscape==T){
				createNetworkFromIgraph(graphInfer, title="Inferred_network",collection=graph.name)
				copyVisualStyle("default","MyStyle01")
				setVisualStyle("MyStyle01")
				setEdgeTargetArrowShapeDefault("Arrow", style.name = "MyStyle01")
				layoutNetwork(layout.name = "cose")

				createNetworkFromIgraph(simplify(endresult), title="Inferred_Motif_network",collection=graph.name)
				setVisualStyle("MyStyle01")
				layoutNetwork(layout.name = "cose")
				
				createNetworkFromIgraph(graph.True, title="True_network",collection=graph.name)
				setVisualStyle("MyStyle01")
				layoutNetwork(layout.name = "cose")
			}
			
			rm(endresult)
		}
		
		setwd('..')	
	}
```



#### Script block 6: Bayesian networks
This code block uses the Bayesian inference method to infer the network data and uses fGetMotifsFromGraph.
```{}
	output.folder <- paste(data_dir,"Bayes",sep="") 
	LinearNet(output.folder, data)
	analyse.output(output.folder)

	prob.file <- paste(output.folder, "/NetworkProbability_Matrix.txt", sep = "")
	prob.mat <- read.table(prob.file)
	prob.mat <- as.matrix(prob.mat)
	threshold = 0.08
	prob.mat[prob.mat < threshold] <- 0

	graph.Bayes <- igraph::graph_from_adjacency_matrix(prob.mat,weighted=TRUE)
	plot(graph.Bayes)
	#nodes: 20
	V(graph.Bayes)
	length(V(graph.Bayes))

	#links: 16
	E(graph.Bayes)
	length(E(graph.Bayes))
	fGetMotifsFromGraph("graph.Bayes",motifsize)
```


#### Script block 7: Regression (Ensemble of trees) networks
This code block uses the Regression based inference method to infer the network data and uses fGetMotifsFromGraph.
```{}
	weightMat <- GENIE3(as.matrix(data))

	linkList <- getLinkList(weightMat)

	linkList <- getLinkList(weightMat, reportMax=5)
	linkList <- getLinkList(weightMat, threshold=0.1)

	graph.Reg <- igraph::graph_from_data_frame(linkList, directed=TRUE, vertices=rownames(data))
	plot(graph.Reg)
	#nodes: 20
	V(graph.Reg)
	length(V(graph.Reg))

	#links: 40
	E(graph.Reg)
	length(E(graph.Reg))
	fGetMotifsFromGraph("graph.Reg",motifsize)
```

#### Script block 8: Mutual information networks
This code block uses the Mutual information based inference method to infer the network data and uses fGetMotifsFromGraph.
```{}
	mi <- knnmi.all(data)
	grn2 <- aracne.m(mi)
	graph.MU2 <- igraph::graph_from_adjacency_matrix(grn2, weighted=TRUE)
	plot(graph.MU2)
	#nodes: 20
	V(graph.MU2)
	length(V(graph.MU2))

	#links: 82
	E(graph.MU2)
	length(E(graph.MU2))
	fGetMotifsFromGraph("graph.MU2",motifsize)
```

#### Post script: Cytoscape
The resulted networks can be imported (and will be imported if runCytoscape=TRUE) into Cytoscape. By evaluating the intersection of the networks, one can assess the prediction accuracy.


